---
title: "p8105_hw6_gl2677"
author: "Gaotong LIU"
date: "11/18/2019"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)# cross validation
library(mgcv) # cross validation

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1
```{r load birth data}
birthweight = read_csv("data/birthweight.csv") %>% 
  mutate(babysex = as.factor(babysex),
         frace  = as.factor(frace),
         malform = as.factor(malform),
         mrace = as.factor(mrace)) %>% 
  mutate(babysex = fct_infreq(babysex),
         frace  = fct_infreq(frace),
         malform = fct_infreq(malform),
         mrace = fct_infreq(mrace))
```

```{r fit regression, include=FALSE}
full_fit = lm(bwt ~ ., data = birthweight)
summary(full_fit)
anova(full_fit)
my_fit = lm(bwt ~ babysex + blength + gaweeks + bhead 
            + delwt + smoken ,
            data = birthweight)
summary(my_fit)
```


```{r  residual plot}
my_fit = lm(bwt ~ babysex + blength + gaweeks + bhead 
            + delwt + smoken ,
            data = birthweight)

birthweight %>% 
  add_residuals(my_fit) %>% 
  add_predictions(my_fit) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  geom_line(aes(y = 0),  color = "gray") +
  geom_smooth(method =  lm, se  = FALSE, linetype = "dashed") + 
  labs(x = "fitted value",
       y = "residual",
       title = "Residuals against fitted values")

```

Firstly I used all predictors to fit the regression, then I remove the predictors which are non-significant. 6 variables show the highest significance and I use them to fit the final regression. The model I use only contains the main effect of the baby’s sex, baby’s length at birth, baby’s head circumference at birth (centimeters), gestational age in weeks, mother’s weight at delivery (pounds) and average number of cigarettes smoked per day during pregnancy.

The residual plot shows the residuals bounce around zero, but there is more variance when the fitted values increase.


```{r cross validation}
cv_df = 
  crossv_mc(birthweight, 5000, test = 0.2) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) %>% 
  mutate(
    fit1_model = map(.x = train, ~lm(bwt ~ blength + gaweeks, 
                                     data = .x)),
    fit2_model = map(.x = train, ~lm(bwt ~ bhead * blength * babysex, 
                                     data = .x)),
    my_fit_model = map(.x = train, ~lm(bwt ~ babysex + blength + gaweeks + bhead + delwt + smoken,
                                     data = .x))
    ) %>% 
    mutate(
    rmse_fit1 = map2(.x = fit1_model, .y = test, ~rmse(.x, .y)),
    rmse_fit2 = map2(.x = fit2_model, .y = test, ~rmse(.x, .y)),
    rmse_my_fit = map2(.x = my_fit_model, .y = test, ~rmse(.x, .y))
  )
```

```{r rmse plot}
cv_df %>%
  select(starts_with("rmse")) %>%
  unnest() %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") %>%
  mutate(model = fct_inorder(model)) %>%
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() + 
  labs(title = "Distribution of root mean squared errors(rmse) against different models")
```

From the distribution of RMSE values for each candidate model, the variance in prediction error of fit1 (`bwt ~ blength + gaweeks`) is larggest, and that of fit2(`bwt ~ bhead * blength * babysex`) and my_fit is quite similar. My fit model `bwt ~ babysex + blength + gaweeks + bhead + delwt + smoken` is the best among three models.

## Problem 2
```{r load weathrer data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r bootstrap}
results = weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    results1 = map(models, broom::glance),
    results2 = map(models, broom::tidy)) %>% 
  mutate(beta_0 = map(.x = results2, ~.x %>% 
                        filter(term =="(Intercept)") %>% 
                        pull(estimate)),
         beta_1 = map(.x = results2, ~.x %>% 
                        filter(term =="tmin") %>% 
                        pull(estimate)),
         r_squared = map(.x = results1, ~.x %>% 
                        pull(r.squared))) %>% 
  select(-strap, -models, -results1, -results2) %>%
  unnest() %>% 
  mutate(log = log(beta_0*beta_1))
```

```{r plot estimate}
results %>% 
  ggplot(aes(x = r_squared)) + 
  geom_density() +
  geom_vline(aes(xintercept = results %>% pull(r_squared) %>% quantile(probs = 0.025)),
             linetype = "dashed") +
  geom_vline(aes(xintercept = results %>% pull(r_squared) %>% quantile(probs = 0.975)),
             linetype = "dashed") +
  labs(title = "Distribution of estimate R squared", 
       subtitle = "with 2.5% and 97.5% quantiles")

results %>% 
  ggplot(aes(x = log)) + 
  geom_density() +
  geom_vline(aes(xintercept = results %>% pull(log) %>% quantile(probs = 0.025)),
             linetype = "dashed") +
  geom_vline(aes(xintercept = results %>% pull(log) %>% quantile(probs = 0.975)),
             linetype = "dashed") +
  labs(title = "Distribution of estimate log(beta_0 * beta_1)", 
       subtitle = "with 2.5% and 97.5% quantiles",
       x = "log(beta_0 * beta_1)")
```

* The distribution of `r_squared` is a nearly normal distribution with slight left skewness. The peak density is obtained when `r_squared` is approximately 0.91, and the range of `r_squared` is from 0.8697 to 0.9431.

* The distribution of `log(beta_0*beta_1)` is a nearly normal distribution and the peak density is obtained when `log(beta_0*beta_1)` is approximately 2.01, and the range of `r_squared` is from 1.918  to 2.106.

* The 95% confidence interval for `r_squared`  is (`r results %>% pull(r_squared) %>% quantile(probs = 0.025) %>% round(digits = 3)`, `r results %>% pull(r_squared) %>% quantile(probs = 0.975) %>% round(digits = 3)`)

* The 95% confidence interval for `log(beta_0*beta_1)`  is (`r results %>% pull(log) %>% quantile(probs = 0.025) %>% round(digits = 3)`, `r results %>% pull(log) %>% quantile(probs = 0.975) %>% round(digits = 3)`)
