---
title: "p8105_hw6_gl2677"
author: "Gaotong LIU"
date: "11/18/2019"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)# cross validation
library(mgcv) # cross validation

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1
```{r load birth data}
birthweight = read_csv("data/birthweight.csv") %>% 
  mutate(babysex = as.factor(babysex),
         frace  = as.factor(frace),
         malform = as.factor(malform),
         mrace = as.factor(mrace))
```

```{r fit regression, include=FALSE}
full_fit = lm(bwt ~ babysex  + blength + gaweeks + 
                   mheight + smoken + wtgain + fincome+ 
                    delwt + malform + parity + 
                bhead +frace + ppbmi + ppwt +
                   menarche + momage + mrace  + 
                pnumlbw + pnumsga ,
            data = birthweight)
anova(full_fit)
initial_fit = lm(bwt ~ babysex + blength + gaweeks + 
                    + bhead + mheight + smoken + wtgain +
                   fincome + delwt + parity +frace + mrace,
            data = birthweight)

my_fit = lm(bwt ~ babysex + blength + gaweeks + bhead,
            data = birthweight)

summary(my_fit)
summary(initial_fit)
summary(full_fit)

```


```{r  residual plot}
my_fit = lm(bwt ~ babysex + blength + gaweeks + bhead,
            data = birthweight)
birthweight %>% 
  add_residuals(my_fit) %>% 
  add_predictions(my_fit) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  geom_line(aes(y = 0),  color = "red")
  
birthweight %>% 
  add_residuals(initial_fit) %>% 
  add_predictions(initial_fit) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  geom_line(aes(y = 0),  color = "red")
```

Firsty I used anava Global F test to find the predictors with significant effect on the birthweight. 12 variables show the significance. Then base on the common sense, and to make the model simpler, I only choose 4 variables which are highly associated with the birthweight. The model I use only contain the main effect of  the baby's sex, baby’s length at birth, baby’s head circumference at birth (centimeters) and gestational age in weeks.

```{r 2 models}
fit1 = lm(bwt ~ blength + gaweeks, data = birthweight)
fit2 = lm(bwt ~ bhead * blength * babysex, data = birthweight)

```

```{r cross validation}
cv_df = 
  crossv_mc(birthweight, 100, test = 0.2) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) %>% 
  mutate(
    fit1_model = map(.x = train, ~lm(bwt ~ blength + gaweeks, 
                                     data = .x)),
    fit2_model = map(.x = train, ~lm(bwt ~ bhead * blength * babysex, 
                                     data = .x)),
    my_fit_model = map(.x = train, ~lm(bwt ~ babysex + blength + gaweeks + bhead,
                                     data = .x))
    ) %>% 
    mutate(
    rmse_fit1 = map2(.x = fit1_model, .y = test, ~rmse(.x, .y)),
    rmse_fit2 = map2(.x = fit2_model, .y = test, ~rmse(.x, .y)),
    rmse_my_fit = map2(.x = my_fit_model, .y = test, ~rmse(.x, .y))
  )
```

```{r rmse plot}
cv_df %>%
  select(starts_with("rmse")) %>%
  unnest() %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") %>%
  mutate(model = fct_inorder(model)) %>%
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

From the distribution of RMSE values for each candidate model. The variance in prediction error of fit1 is larggest, and that of fit2 and my_fit is quite similar. My fit model `bwt ~ babysex + blength + gaweeks + bhead` is the best among three models.

## Problem 2
```{r load weathrer data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r bootstrap}
results = weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    results1 = map(models, broom::glance),
    results2 = map(models, broom::tidy)) %>% 
  mutate(beta_0 = map(.x = results2, ~.x %>% 
                        filter(term =="(Intercept)") %>% 
                        pull(estimate)),
         beta_1 = map(.x = results2, ~.x %>% 
                        filter(term =="tmin") %>% 
                        pull(estimate)),
         r_squared = map(.x = results1, ~.x %>% 
                        pull(r.squared))) %>% 
  select(-strap, -models, -results1, -results2) %>%
  unnest() %>% 
  mutate(log = log(beta_0*beta_1))
```
